---
title: "Recreating Dr. Zhang's Analysis"
---
```{r}

library(lme4)
library(dplyr)
library(magrittr)

library(lmerTest)
library(interactions)

library(emmeans)    

library(nlme)
library(performance) # for calculating ICC

```

# (0) Loading the data

```{r}
Data <- read.csv("/Users/ekindogailkel/Library/Mobile Documents/com~apple~CloudDocs/PI_manuscript/Regression_Hui/newstruct1.csv")

nrow(unique(Data["subj"])) # 105
nrow(Data["subj"]) # 14055
# the Data is already in the long format 
```


```{r}

# Change continuous data to categorical data 

Data$subj<-factor(Data$subj)
Data$gene<-factor(Data$gene)
Data$condi<-factor(Data$condi)
Data$gender<-factor(Data$gender)
Data$age_category<-factor(Data$age_category)

```


# (1) Null model:  

Drop error = task performance, subj = grouping variable 

```{r}
null_model <- lmer(dropError~1+(1|subj), REML=FALSE, data = Data)
summary(null_model)
```
# (1.b) Calculating ICC:  

```{r}
ICC <- icc(null_model)

print(ICC)
```

# (1.c) Calculating DEFF

DEFF is a term used for deciding whether grouping is needed.

Kish, 1965, p.162.

DEFF = 1 + ICC(no_observations/no_clusters - 1)

```{r}
DEFF <- 1 + 0.25*(nrow(Data["subj"])/nrow(unique(Data["subj"])) - 1)

print(DEFF)
```


# (2) Decomposition of PI performance (dropError) 

- modeling dropError as a function of angleError and distanceError
- random intercepts and random slopes for the grouping variable subj

angleError = angle between the ideal and actual inbound path 
distanceError = distance between the ideal and actual inbound path 


```{r}
dropError_decomp <- lmer(dropError ~ angleError + distanceError + (1+1|subj), REML = FALSE, data = Data)
summary(dropError_decomp)
confint(dropError_decomp)
```


# (3) Various factors impacting PI performance: (gene*gender*condi*age)

```{r}
dropError_all <- lmer(dropError ~ gene*gender*condi*age +(1+1|subj), REML = FALSE, data = Data)
summary(dropError_all)     
confint(dropError_all)
```

## Simple (Gene, gender):

gender1 = male, gene1 = carrier
```{r}
emm1 <- emmeans(dropError_all, ~ gene : gender)
pairs(emm1, simple = "each")
```


# (4) Various factors impacting PI performance: (gene*gender*condi*age category)

```{r}
dropError_all2 <- lmer(dropError~ gene*gender*condi*age_category + (1+1|subj), REML = FALSE, data = Data)
summary(dropError_all2)
confint(dropError_all2)
```


```{r}
emm3 <- emmeans(dropError_all2, ~ age_category:gene)
pairs(emm3, simple = "each")
```
```{r}
emm4 <- emmeans(dropError_all2, ~ age_category:gene:gender)
pairs(emm4, simple = "each")
```


# 5 Mediation analysis1: gender/age/gene - distance error - drop error 
# (5.a) Mediation analysis1: Interaction effects on distance error

```{r}
media1_step2 <- lmer(distanceError~ gene*gender*condi*age +  (1+1|subj), REML = FALSE, data = Data)
summary(media1_step2)   
confint(media1_step2)
```
# (5.b) Mediation analysis1: Distance error on PI (drop error)

```{r}
media1_step3 <- lmer(dropError~ distanceError + (1+1|subj), REML = FALSE, data =  Data)  
summary(media1_step3)
confint(media1_step3)
```

# (5.c) Mediation analysis1: Distance error + interactions on PI (drop error)

```{r}
media1_step4 <- lmer(dropError~ gene*gender*condi*age + distanceError + (1+1|subj), REML = FALSE, data = Data)
summary(media1_step4)
confint(media1_step4)
```


# 6 Mediation analysis2: gender/age/gene - angle error - drop error 
# (6.a) Mediation analysis2: Interaction effects on angle error

```{r}
media2_step2 <- lmer(angleError~ gene*gender*condi*age +
                      (1+1|subj), REML = FALSE, data = Data)
summary(media2_step2)  
confint(media2_step2)
```

# (6.b) Mediation analysis2: Angle error on PI (drop error)

```{r}
media_step2_3 <- lmer(dropError~ angleError + (1+1|subj), REML = FALSE, data = Data)
summary(media_step2_3)
confint(media_step2_3)
```

# (6.c) Mediation analysis2: Angle error + interactions on PI (drop error)

```{r}
media_step2_4 <- lmer(dropError~ gene*gender*condi*age + angleError + (1+1|subj), REML = FALSE, data = Data)
summary(media_step2_4)
confint(media_step2_4)                      
```

# (7) Regressing out both angle error and distance error

```{r}
regress_all <- lmer(dropError~ gene*gender*condi*age + angleError +  distanceError + (1+1|subj), REML = FALSE, data = Data)
summary(regress_all)
confint(regress_all)
```

# (8) Experimental manipulation (condi) - elapsedSearchTime
```{r}
cond_1 <- lmer(elapsedSearchTime ~ condi+  (1+condi|subj), REML = FALSE, data = Data)
summary(cond_1)
confint(cond_1)
```
# (9)  Experimental manipulation (condi) - elapsedSearchTime - Excluding non moving periods ?

# (10) Experimental manipulation (condi) - elapsed path length 
```{r}
cond_2 <- lmer(actural_path_lenth ~ condi+  (1+condi|subj), REML = FALSE, data = Data)
summary(cond_2)
confint(cond_2)
```          

# (11) Experimental manipulation (condi) - navigation trajectory 

```{r}
cond_3 <- lmer(dis_to_actural ~ condi+  (1+condi|subj), REML = FALSE, data = Data)
summary(cond_3)
confint(cond_3)
```         


# Subsetting the data as - teleportation and continuous 
condition = 1 -> teleport

```{r}
Data_teleport <- subset(Data, condi == 1)
Data_cont <- subset(Data, condi == 0)
```


# (12) Teleportation condition: Elapsed searching time on drop error (significant) 
```{r}
model_teleport1 <- lmer(dropError ~ elapsedSearchTime + (1|subj), REML = FALSE, data = Data_teleport)
summary(model_teleport1)  
confint(model_teleport1)
```

# (13) Continuous condition: Elapsed searching time on drop error (NOT-significant) 

```{r}
model_cont1 <- lmer(dropError ~ elapsedSearchTime + (1|subj), REML = FALSE, data = Data_cont)
summary(model_cont1)      
confint(model_cont1)
```

# (14) Continuous condition: path length on drop error (significant)


```{r}
model_cont2 <- lmer(dropError ~ actural_path_lenth + (1|subj), REML = FALSE, data = Data_cont)
summary(model_cont2)       
confint(model_cont2)
```

# (15) Teleportation condition: path length on drop error (NOT-significant)


```{r}
model_teleport2 <- lmer(dropError ~ actural_path_lenth + (1|subj), REML = FALSE, data = Data_teleport)
summary(model_teleport2)     
confint(model_teleport2)
```

# 16 Mediation analysis3: Continuous: path length - distance error - drop error 

# (16.a) Mediation analysis3: path length on distance error (step2)
```{r}
model_cont3 <- lmer(distanceError ~ actural_path_lenth + (1|subj), REML = FALSE, data = Data_cont)
summary(model_cont3)  
confint(model_cont3)
```

# (16.b) Mediation analysis3: distance error on drop error (step3)

```{r}
model_cont4 <- lmer(dropError ~ distanceError + (1|subj), REML = FALSE, data = Data_cont)
summary(model_cont4)  
confint(model_cont4)
```


# (16.c) Mediation analysis3: distance error + path length on drop error (step4)

```{r}
model_cont5 <- lmer(dropError ~ actural_path_lenth + distanceError + (1|subj), REML = FALSE, data = Data_cont)
summary(model_cont5)  
confint(model_cont5)
```


# 17 Mediation analysis4: Continuous - path length - angle error - drop error 
# (17.a) Mediation analysis4:path length on angle error (step2)

```{r}
model_cont6 <- lmer(angleError ~ actural_path_lenth + (1|subj), REML = FALSE, data = Data_cont)
summary(model_cont6)  
confint(model_cont6)
```

# (17.b) Mediation analysis4: angle error on drop error (step3)
```{r}
model_cont7 <- lmer(dropError ~ angleError + (1|subj), REML = FALSE, data = Data_cont)
summary(model_cont7)  
confint(model_cont7)
```


# (17.c) Mediation analysis4: angle error + path length on drop error (step4)
```{r}
model_cont8 <- lmer(dropError ~ actural_path_lenth + angleError + (1|subj), REML = FALSE, data = Data_cont)
summary(model_cont8)  
confint(model_cont8)
```



# 18 Mediation analysis5: Teleportation: searching time - angle error - drop error 
# (18.a) Mediation analysis5: searching time on angle error (step2) - NOT-significant







```{r}
model_teleport3 <- lmer(angleError ~ elapsedSearchTime + (1|subj), REML = FALSE, data = Data_teleport)
summary(model_teleport3)  
```

# (18.b) Mediation analysis5: angle error on drop error 

```{r}
model_teleport4 <- lmer(dropError ~ angleError + (1|subj), REML = FALSE, data = Data_teleport)
summary(model_teleport4)  
confint(model_teleport4)
```

# (18.c) Mediation analysis5: angle error + search time on drop error


```{r}
model_teleport7 <- lmer(dropError ~ elapsedSearchTime + angleError + (1|subj), REML = FALSE, data = Data_teleport)
summary(model_teleport7)
confint(model_teleport7)
```



# 19 Mediation analysis6: Teleportation - searching time - distance error - drop error 
# (19.a) Mediation analysis6: searching time on distance error (not signicant)

```{r}
model_teleport8 <- lmer(distanceError ~ elapsedSearchTime + (1|subj), REML = FALSE, data = Data_teleport)
summary(model_teleport8)    
confint(model_teleport8)
```

# (20) Gene on signed distance error
```{r}
signed_distance_model <- lmer(signed_distanceError ~ gene + (1+1|subj), REML = FALSE, data = Data)
summary(signed_distance_model)
confint(signed_distance_model)
```

# (21) Gene on angle error 

```{r}
angle_error_model <- lmer(angleError ~ gene + (1+1|subj), REML = FALSE, data = Data)
summary(angle_error_model)
confint(angle_error_model)
```



